---
title: "Gapminder dataset example"
author: Melissa Cheung
output: html_document
date: "2022-12-01"
---

```{r setup, message=FALSE}
library(tidyverse)
library(rmarkdown)
library(kableExtra)
library(ggpubr)
```


```{r load-data, message=FALSE}

#read file into R
pop_data <- read_csv("data/gapminder_data.csv")

#create a table with data from European countries in 2007 showing the countries 
#with the largest life expectancy at the top 
euro_data_tbl <- pop_data %>% 
                    filter(continent == "Europe" & year == 2007) %>% 
                    select(-continent, -year) %>% 
                    arrange(desc(lifeExp)) %>% 
                    rename(Country = country, "Population Size" = pop, 
                      "Life Expectancy" = lifeExp, "GDP" = gdpPercap)
    
```

```{r kbl}
euro_data_tbl %>% 
          kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive", full_width = F))
```

## Paged table
```{r paged-table}
paged_table(euro_data_tbl)
```

## Challenge
Create a new dataset euro_data_fig by filtering the pop_data tibble to contain only data from Europe. Draw a plot to display the lifeExp on the y axis and year on the x axis. Use geom_violin() to draw this as a violin plot to show the distrubution of the data across each year and save it in a euro_plot variable.

```{r euro-data}
euro_data_fig <- pop_data %>%
  filter(continent == "Europe") %>% 
                    mutate(year=as_factor(year))

euro_plot <- ggplot(data = euro_data_fig, 
                    aes(x = year, y = lifeExp))+
  geom_violin()
euro_plot
```

Create a new dataset uk_data_fig by filtering the pop_data tibble to contain only data from the United Kingdom. Draw a scatter plot to display the lifeExp on the y axis and year on the x axis and save it in a uk_plot variable.
Draw the euro_plot created in the previous challenge next to a uk_plot using the ggarrange() function. Label the plots A and B respectively.
```{r uk-data}
uk_data_fig <- pop_data %>%
  filter(country == "United Kingdom")

uk_plot <- ggplot(data = uk_data_fig, aes(x = year, y = lifeExp))+
  geom_point()

uk_plot

ggarrange(euro_plot, uk_plot,
          labels = c("A", "B"))
```










